A SAT-Based Counterexample Guided Method for Unbounded Synthesis

### 2016-04-22 paper.tex
Refer to the example in section 2

First, we introduce a running example to assist the explanation. We consider a
simple arbiter system in which the environment requests for either 1 or 0
resources, and the controller may grant access to two resources. The controller
must ensure that the number of unhandled requests is not $\geq 2$.
Figure~\ref{fig:example} shows the variables (\ref{fig:examplevars}) and the
formulas for computing next-state variables (\ref{fig:exampletrans}) for this
example.

\begin{figure}
    \begin{subfigure}[t]{\textwidth}
        \centering
        \begin{tabular}{l | l | l}
            \textbf{Controllable} & \textbf{Uncontrollable} & \textbf{State} \\
            \hline
            \texttt{request = \{0, 1\}} & \texttt{grant0 = \{0, 1\}} & \texttt{resource0 = \{0, 1\}} \\
            & \texttt{grant1 = \{0, 1\}} & \texttt{resource1 = \{0, 1\}} \\
            & & \texttt{nrequests = \{0, 1, 2, 3\}} \\
        \end{tabular}
        \caption{Variables}
        \label{fig:examplevars}
    \end{subfigure}

@@

The algorithm consists of two competing solvers, one for the controller and one for
the environment, that solve abstractions of the safety game. Each solver builds
a candidate strategy for its corresponding player consisting of assignments to
$2^C$ and $2^U$ respectively.

@@

An abstraction of the game restricts actions available to one of the players.
Specifically, we consider abstractions represented as trees of actions,
referred to as \emph{abstract game trees} (AGTs).  Figure~\ref{fig:agt} shows
an example abstract game tree restricting the controller (abstract game trees
restricting the environment are similar).  In the abstract game, the controller
is required to pick actions from the tree, starting from the root node.  After
reaching a leaf, it continues playing unrestricted.  The tree in
Figure~\ref{fig:agt} restricts the initial controller action to the set $\{a,
d\}$.  After choosing action $a$, the controller is required to play action $b$
in the next round.  It then reaches a leaf of the tree and continues playing
unrestricted.  The root of the tree is annotated by the initial state $s$ of
the abstract game and the bound $k$ on the number of rounds.  We denote
$\textsc{nodes}(T)$ the set of all nodes of a tree $T$, $\textsc{leaves}(T)$
the subset of leaf nodes.  For edge $e$, $\textsc{action}(e)$ is the action
that labels the edge, and for node $n$, $\textsc{height}(k, n)$ is the distance
from n to the last round of a game bounded to $k$ rounds.  $\textsc{height}(k,
T)$ is the height of the root node of the tree.  For node $n$ of the tree,
$\textsc{succ}(n)$ is the set of pairs $\langle e, n' \rangle$ where $n'$ is a
child node of $n$ and $e$ is the edge connecting $n$ and $n'$.

@@

                        edge from parent [fixed] node [left] {\texttt{req=0}}

@@

                        node [left=4pt] {\texttt{gr0=0, gr1=0}}
                        edge from parent [fixed] node [left] {\texttt{req=0}}

@@

                    node [left=4pt] {\texttt{gr0=0, gr1=0}}
                    edge from parent [fixed] node [left] {\texttt{req=0}}

@@

                        edge from parent [fixed] node [right] {\texttt{req=1}}

@@

Given a controller (environment) abstract game tree $T$ a \emph{partial
strategy} $Strat: \textsc{nodes}(T) \rightarrow U$ ($Strat: \textsc{nodes}(T)
\rightarrow C$) labels each node of the tree with the opponent's action to be
played in that node.   Figure~\ref{fig:strategy} shows an example partial
strategy.  The environment starts by choosing action $\alpha$.  If the
controller plays $a$, the environment responds with $\beta$ in the next round,
and so on. Given a partial strategy $Strat$, we can map each leaf $l$ of the
abstract game tree to $\langle s',i'\rangle=\textsc{outcome}(\langle s,
i\rangle, Strat, l)$ obtained by playing all controllable and uncontrollable
actions on the path from the root to the leaf.

@@

If the dual solver can find no spoiling strategy at any of the leaves, then the
candidate partial strategy is a winning one. Otherwise, \textsc{verify} returns
the move used by the opponent to defeat a leaf of the partial strategy, which
is appended to the corresponding node in $T$ in order to refine it in line~(9)
as shown in Figure~(\ref{fig:refined}).

>>>>

First, we introduce a running example to assist the explanation. We consider a
simple arbiter system in which the environment requests for either 1 or 2
resources, and the controller may grant access to two resources. The controller
must ensure that the number of unhandled requests is not $\geq 3$.
Figure~\ref{fig:example} shows the variables (\ref{fig:examplevars}) and the
formulas for computing next-state variables (\ref{fig:exampletrans}) for this
example.

\begin{figure}
    \begin{subfigure}[t]{\textwidth}
        \centering
        \begin{tabular}{l | l | l}
            \textbf{Controllable} & \textbf{Uncontrollable} & \textbf{State} \\
            \hline
            \texttt{request = \{1, 2\}} & \texttt{grant0 = \{0, 1\}} & \texttt{resource0 = \{0, 1\}} \\
            & \texttt{grant1 = \{0, 1\}} & \texttt{resource1 = \{0, 1\}} \\
            & & \texttt{nrequests = \{0, 1, 2, 3\}} \\
        \end{tabular}
        \caption{Variables}
        \label{fig:examplevars}
    \end{subfigure}

@@

The bounded synthesis algorithm consists of two competing solvers, one for the
controller and one for the environment, that solve abstractions of the safety
game. Each solver builds a candidate strategy for its corresponding player
consisting of assignments to $2^C$ and $2^U$ respectively.

@@

An abstraction of the game restricts actions available to one of the players.
Specifically, we consider abstractions represented as trees of actions,
referred to as \emph{abstract game trees} (AGTs).  Figure~\ref{fig:agt} shows
an example abstract game tree restricting the environment (abstract game trees
restricting the controller are similar).  In the abstract game, the controller
can freely choose actions whilst the environment is required to pick actions
from the tree.  After reaching a leaf, the enivironment continues playing
unrestricted.  The tree in Figure~\ref{fig:agt} restricts the first and second
environment actions to \texttt{request=1}. It then reaches a leaf of the tree
and continues playing unrestricted.

The root of the tree is annotated by the initial state $s$ of the abstract game
and the bound $k$ on the number of rounds.  We denote $\textsc{nodes}(T)$ the
set of all nodes of a tree $T$, $\textsc{leaves}(T)$ the subset of leaf nodes.
For edge $e$, $\textsc{action}(e)$ is the action that labels the edge, and for
node $n$, $\textsc{height}(k, n)$ is the distance from n to the last round of a
game bounded to $k$ rounds.  $\textsc{height}(k, T)$ is the height of the root
node of the tree.  For node $n$ of the tree, $\textsc{succ}(n)$ is the set of
pairs $\langle e, n' \rangle$ where $n'$ is a child node of $n$ and $e$ is the
edge connecting $n$ and $n'$.

@@

                        edge from parent [fixed] node [left] {\texttt{req=1}}

@@

                        node [left=4pt] {\texttt{gr0=1, gr1=0}}
                        edge from parent [fixed] node [left] {\texttt{req=1}}

@@

                    node [left=4pt] {\texttt{gr0=0, gr1=1}}
                    edge from parent [fixed] node [left] {\texttt{req=1}}

@@

                        edge from parent [fixed] node [right] {\texttt{req=2}}

@@

Given an environment (controller) abstract game tree $T$ a \emph{partial
strategy} $Strat: \textsc{nodes}(T) \rightarrow C$ ($Strat: \textsc{nodes}(T)
\rightarrow U$) labels each node of the tree with the controller's
(environment's) action to be played in that node.   Given a partial strategy
$Strat$, we can map each leaf $l$ of the abstract game tree to $\langle
s',i'\rangle=\textsc{outcome}(\langle s, i\rangle, Strat, l)$ obtained by
playing all controllable and uncontrollable actions on the path from the root
to the leaf.  Figure~\ref{fig:strategy} shows an example partial strategy for
the controller.  The controller responds to the requests by first granting
access to \texttt{resource1}, then to \texttt{resource0} in the second round.

@@

If the dual solver can find no spoiling strategy at any of the leaves, then the
candidate partial strategy is a winning one. Otherwise, \textsc{verify} returns
the move used by the opponent to defeat a leaf of the partial strategy, which
is appended to the corresponding node in $T$ in order to refine it in line~(9).
In our example, the controller's partial strategy in Figure~\ref{fig:strategy}
is defeated by the counterexample action \texttt{req=2}.
Figure~\ref{fig:refined} shows a refined AGT that includes this action.
